---
Task-1: Prediction using Unsupervised Learning
Name: Vinamrata Rani
---

K-Means Clustering Analysis on Iris Dataset




#Installing *packages*

```{r}
install.packages("tidyverse")


install.packages("cluster")


install.packages("reshape2")
```



#Loading *packages*
```{r}
library(ggplot2)
library(cluster)
library(tidyverse)
library(reshape2)
```


#Import the Iris Data set
```{r}
my_data = read.csv("iris.csv")
glimpse(my_data)
head(my_data)
View(my_data)

```



#Explore the Data set
```{r}
class(my_data)
str(my_data)
```


#Visualizing the data with *ggplot2*

*SepalLength vs SepalWidth*

```{r}
ggplot(my_data, aes(SepalLengthCm, SepalWidthCm)) + geom_point() + ggtitle(" Iris scatter plot (SepalLength vs SepalWidth)")
```


*PetalLength vs PetalWidth*
```{r}
##library(ggplot2)
ggplot(my_data, aes(PetalLengthCm, PetalWidthCm)) + geom_point() + ggtitle(" Iris scatter plot (PetalLength vs PetalWidth)")
```




*SepalLength vs PetalLength*

```{r}
ggplot(my_data, aes(SepalLengthCm, PetalLengthCm)) + geom_point() + ggtitle(" Iris scatter plot (SepalLength vs PetalLength)")
```





*SepalWidth vs PetalWidth*

```{r}
ggplot(my_data, aes(SepalWidthCm, PetalWidthCm)) + geom_point() + ggtitle(" Iris scatter plot (SepalWidth vs PetalWidth)")
```





#Box plots

```{r}
ggplot(my_data)+
 geom_boxplot(aes(x = Species, y = SepalLengthCm, fill = Species))+
 theme_bw()
ggplot(my_data)+
 geom_boxplot(aes(x = Species, y = SepalWidthCm, fill = Species))+
 theme_bw()
ggplot(my_data)+
 geom_boxplot(aes(x = Species, y = PetalLengthCm, fill = Species))+
 theme_bw()
ggplot(my_data)+
 geom_boxplot(aes(x = Species, y = PetalWidthCm, fill = Species))+
 theme_bw()
```





#Find the optimal number of clusters by Elbow Method


```{r}
set.seed(150) # for reproduction
wcss <- vector()
for (i in 1:10) wcss[i] <- sum(kmeans(my_data[, 2:5], i)$withinss)
plot(1:10,
 wcss,
 type = "b",
 main = paste("The Elbow Method"),
 xlab = "Number of Clusters",
 ylab = "WCSS"
)
```
the elbow point : *k(centers) = 3*






#Apply kmeans function to the feature columns


```{r}
set.seed(150)

cluster.iris <- kmeans(my_data[, 2:5],centers =  3)
cluster.iris
```
#The kmeans has grouped the data into three clusters- 1, 2 & 3 having *50, 62 & 38* observations respectively

```{r}
yclus <- cluster.iris$cluster
table(yclus)
```


#Visualize the kmeans clusters


```{r}
clusplot(my_data[, -5],
 yclus,
 lines = 0,
 shade = TRUE,
 color = TRUE,
 labels = 0,
 plotchar = FALSE,
 span = TRUE,
 main = paste("Clusters of Iris Flowers")
)
```


#Compare the clusters


```{r}
my_data$cluster.kmean <- yclus

c_table=table(my_data$Species , my_data$cluster.kmean)
c_table
```





#Tiles plot : Species vs. kmeans clusters

```{r}
mtable <- melt(c_table)
ggplot(mtable)+
 geom_tile(aes(x = Var1, y = Var2, fill = value))+
 labs(x = "Species", y = "kmeans Cluster")+
 theme_bw()
```



#Scatter plots (to view Species & kmeans custers)


*SepalLength vs SepalWidth*

```{r}
cluster.iris$cluster <- as.factor(cluster.iris$cluster)
ggplot(my_data, aes(SepalLengthCm, SepalWidthCm, color=Species)) + geom_point() + ggtitle(" Iris scatter plot (SepalLength vs SepalWidth) - Species") + theme_bw()


ggplot(my_data, aes(SepalLengthCm, SepalWidthCm, color=cluster.iris$cluster)) + geom_point() + ggtitle("Iris  Kmeans cluster plot") + theme_bw()

```



*PetalLength vs PetalWidth*
```{r}
ggplot(my_data) + geom_point(aes(PetalLengthCm, PetalWidthCm, col=Species)) + ggtitle(" Iris scatter plot (PetalLength vs PetalWidth)")


ggplot(my_data) + geom_point(aes(PetalLengthCm, PetalWidthCm, col=cluster.iris$cluster)) + ggtitle("Iris  Kmeans cluster plot") 

```


*SepalLength vs PetalLength*
```{r}
ggplot(my_data, aes(SepalLengthCm, PetalLengthCm, col=Species)) + geom_point() + ggtitle(" Iris scatter plot (SepalLength vs PetalLength)")

ggplot(my_data,aes(SepalLengthCm, PetalLengthCm, col=cluster.iris$cluster)) + geom_point() + ggtitle("Iris  Kmeans cluster plot")
```




*SepalWidth vs PetalWidth*
```{r}
ggplot(my_data, aes(SepalWidthCm, PetalWidthCm, color=Species)) + geom_point() + ggtitle(" Iris scatter plot (SepalWidth vs PetalWidth)")

ggplot(my_data, aes(SepalWidthCm, PetalWidthCm, color=cluster.iris$cluster)) + geom_point() + ggtitle(" Iris  Kmeans cluster plot")
```




